Replicating Blair and Sambanis: Forecasting Civil Wars - JCR 2020

### 2020-04-29 Tables/Table 1.Rmd
removed r time stamp
Created new figure 1 and table 1
shows smoothed and not smoothed ROC plots and AUC-ROC scores for the base specification

### Introduction^[RM: As a way to maintain transparency and as a poor-mans version of preregistration, I'm pushing each stepto GitHub, where it will have a time-stamp and track all changes.]

One issue that stood out as we read through Blair and Sambanis' (2020) article was how smooth their Receiver Operating Characteristic curves where in Figure 1 on page 14. A Receiver Operating Characteristic curve displays the balance between the true positive rate and the false positive rate, across a range of acceptance thresholds -- the value in which a zero to one probability is assumed to be an "onset" (here: civil conflict). Given that the DV is civil conflict onset (1/0) and that this is a rare event, these curves should resemble a step-like function. Looking through the Blair/Sambanis replication code, we noticed that they we using a rarely used in practice smoothing function when producing their Receiver Operating Characteristic curves.

@@

Note: we are only reproducing the "Base specification" scores found in Table 1, as these are the what Blair and Sambanis use to support their calims. In order to ensure that this replication is a close as possible to the authors' original work, I will be using the replication code they provided at https://journals.sagepub.com/doi/suppl/10.1177/0022002720918923. There is a lot going on in their code, so I will be commenting out all elements that are not necessary for this replication process. Futher, I will note any changes that I make this parsed code (i.e., removing the smoothing function). The raw replication files can be found at `GitHub/Blair-Sambanis-replication/Replication files/replication-3`, while all my edited code can be found at `GitHub/Blair-Sambanis-replication/Replication files/replication-3_BM`

```{r}
print(Sys.time())
```

>>>>

### Introduction^[RM: As a way to maintain transparency and as a poor-mans version of preregistration, I'm pushing each step to GitHub, where it will have a time-stamp and track all changes.]

One issue that stood out as we read through Blair and Sambanis' (2020) article was how smooth their Receiver Operating Characteristic curves where in Figure 1 on page 14. A Receiver Operating Characteristic curve displays the balance between the true positive rate and the false positive rate, across a range of discrimination thresholds -- the value in which a zero to one probability is assumed to be an "onset" (here: civil conflict). Given that the DV is civil conflict onset (1/0) and that this is a rare event, these curves should resemble a step-like function. Looking through the Blair/Sambanis replication code, we noticed that they we using a rarely used in practice smoothing function when producing their Receiver Operating Characteristic curves.

@@

Note: we are only reproducing the "Base specification" scores found in Table 1, as these are the what Blair and Sambanis use to support their claims. In order to ensure that this replication is a close as possible to the authors' original work, I will be using the replication code they provided at https://journals.sagepub.com/doi/suppl/10.1177/0022002720918923. There is a lot going on in their code, so I will be commenting out all elements that are not necessary for this replication process. Further, I will note any changes that I make this parsed code (i.e., removing the smoothing function). The raw replication files can be found at `GitHub/Blair-Sambanis-replication/Replication files/replication-3`, while all my edited code can be found at `GitHub/Blair-Sambanis-replication/Replication files/replication-3_BM`

### 2020-04-29 issues.Rmd
Created new figure 1 and table 1
shows smoothed and not smoothed ROC plots and AUC-ROC scores for the base specification

While reading Blair and Sambanis (2020), a few issues stuck out. We downloaded the replication files authors provide from https://journals.sagepub.com/doi/suppl/10.1177/0022002720918923/suppl_file/sj-zip-1-jcr-10.1177_0022002720918923.zip. The original folder name of is `sj-zip-1-jcr-10.1177_0022002720918923/replication-3`. For simplicity, we extracted this `replication-3` folder and saved to the `Replication Files` folder of this repo.

@@

- Improperly tuning of models
  - RM: In a twitter DM, I asked Blair to clarify -- "[I] was wondering how you all came to your tuning procedure for the RF model hyper-parameters?" He responded "mostly trial and error, honestly. ... Trial and error was on early data. Forecasts were for much later data"
- Using a regression framework in a classification problem
- Their train/test approach 
  - 5-year forecasts for sep plots and AUC scores)
- Rounding of AUC/ROC scores
- Lack of yearly test forecasts in favor of a single 5-year test forecast
- Table 4! 
  - There are only 20 onsets from 2001 to 2015 But there are 15 "Persistence" cases (top of table 4) and 16 "Change" cases (bottom of table 4)
- Treating the output of a RF regression model as Pr() -- (RM: I guess this is akin to a Linear Probability Model, but I'm not sure if this is possible in a RF framework
- The lack of procedures to account for rare events in an RF model

>>>>

While reading Blair and Sambanis (2020), a few issues stuck out. We downloaded the replication files authors provide from https://journals.sagepub.com/doi/suppl/10.1177/0022002720918923/suppl_file/sj-zip-1-jcr-10.1177_0022002720918923.zip. The original folder name of is `sj-zip-1-jcr-10.1177_0022002720918923/replication-3`. For simplicity, we extracted this `replication-3` folder and saved to the `Replication Files` folder of this repo. To ensure that we have a clean copy of the original replication scripts, I copied all files to `replication-3_BM` I'm (RM) currently using the sripts in this folder to recreate figure 1 and table 1 with and without smoothing -- see below

@@

- Improperly tuning of models
  - RM: In a twitter DM, I asked Blair to clarify -- "[I] was wondering how you all came to your tuning procedure for the RF model hyper-parameters?" He responded "mostly trial and error, honestly. ... Trial and error was on early data. Forecasts were for much later data"
- Using a regression framework in a classification problem
- Their train/test approach 
  - 5-year forecasts for sep plots and AUC scores)
- Smoothing of AUC/ROC scores
- Lack of yearly test forecasts in favor of a single 5-year test forecast
- Table 4! 
  - There are only 20 onsets from 2001 to 2015 But there are 15 "Persistence" cases (top of table 4) and 16 "Change" cases (bottom of table 4)
- Treating the output of a RF regression model as Pr() -- (RM: I guess this is akin to a Linear Probability Model, but I'm not sure if this is possible in a RF framework
- The lack of procedures to account for rare events in an RF model
